---
title: "BIOL606 WED 4/19/23 HW LEC16 Cho HTML NOTEBOOK"
author: "Eddie Cho"
due date: 04/19/2023
output:
  html_document:
    df_print: paged
---

#Step 0: Verify current directory and verify the data file 
```{r}
print("The analysis started...")

cat("\n")

print(paste("YOUR current working directory is:", getwd()))

cat("\n")

print(paste("Verifying if the data file for this HW is in YOUR working directory...'True' if it exists, 'False' if it does NOT exists:", file.exists('macnally_full2.csv')))
```

#Step 1: Invoke necessary libraries/packages
```{r setup, message=FALSE}

library(vegan)
library(tidyr)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(ggfortify)

print("Done. All packages loaded")
```

#Step 2: Import data
```{r}
bird <- read.csv("macnally_full2.csv", header=T)

print("Done. Data loaded and saved as dataframe called 'bird'. ")

head(bird) #Display first 6 rows

dim(bird) #Display Dimension
```

#Step 3-A (HW Question #1 / Prompt #1): Run a NMDS on the bird data (macnally_full2.csv, leave out columns 1&2, SITE + HABITAT), using Bray-Curtis as the distance metric with the vegan metaMDS function
```{r}
#Remove 1st column (SITE, -1)  and 2nd column (HABITAT, -2) and save a new data frame, 'new_bird'
new_bird <- bird[,c(-1, -2)]

head(new_bird) #Display first 6 rows

dim(new_bird) #Display Dimension
```

#Step 3-B (HW Question #1 / Prompt #1): Run a NMDS on the bird data (macnally_full2.csv, leave out columns 1&2, SITE + HABITAT), using Bray-Curtis as the distance metric with the vegan metaMDS functions
```{r}
##NOTE to SELF: metaMDS function likes numerical values, hence use the new data frame with removed columns, SITE & HABITAT 

#Use Bray-Curtis as the distance metric with the vegan metaMDS functions
bc.nmds <- metaMDS(new_bird, distance = "bray", k=2, trace=0)

#print bc.nmds
print(bc.nmds)
```

#Step 4-A (HW Question #2 / Prompt #2): Generate a NMDS ordination plot showing the sites only (no species arrows), and draw ellipsoid hulls around each of the 6 habitat types to show how habitat maps onto the ordination.
```{r}
#create empty plot
nmds_ordiplot = ordiplot(bc.nmds, type='none', main = paste0('Stress = ', round(bc.nmds$stress, dig=3) ), las =1 )

#Then, create polygons based on HABITAT using ordihull
ordihull(bc.nmds,  draw ='polygon', groups = bird$HABITAT, label = T)
```

#Step 4-B (HW Question #2 / Prompt #2): Color the site points by HABITAT type (column 2 of data frame)
```{r}
#create empty plot
nmds_ordiplot = ordiplot(bc.nmds, type='none', main = paste0('Stress = ', round(bc.nmds$stress, dig=3) ), las =1 )

#Then, assign 6 colors based on 6 HABITATS using ordihull
ordihull(bc.nmds,  draw ='polygon', groups = bird$HABITAT, col=c('skyblue','orange', 'yellow', 'black', 'green', 'purple' ), label = T)
```

#Step 4-C (HW Question #3 / Prompt #3): Use the ordiellipse() function, and the groups argument should be defined using the HABITAT data in column 2 of the data frame
```{r}
#create empty plot
nmds_ordiplot = ordiplot(bc.nmds, type='none', main = paste0('Stress = ', round(bc.nmds$stress, dig=3) ), las =1 )

#Then, assign 6 colors based on 6 HABITATS using ordiellipse
ordiellipse(bc.nmds,  draw ='polygon', groups = bird$HABITAT, col=c('skyblue','orange', 'yellow', 'black', 'green', 'purple' ), label = T)

```

#Step 4-C (HW Question #3 / Prompt #3): Use the ordiellipse() function, and the groups argument should be defined using the HABITAT data in column 2 of the data frame
```{r}
#create empty plot
nmds_ordiplot = ordiplot(bc.nmds, type='none', main = paste0('Stress = ', round(bc.nmds$stress, dig=3) ), las =1 )

#Then, assign 6 colors based on 6 HABITATS using ordie llipse
ordiellipse(bc.nmds,  draw ='polygon', groups = bird$HABITAT, col=c('skyblue','orange', 'yellow', 'black', 'green', 'purple' ), label = F)

#In addition, add legends based on assinged 6 colors
legend('topleft', legend=c('Mixed', 'Gipps.Manna', 'Montane Forest', 'Foothills Woodland', 'Box-Ironbark', 'River Red Gum' ), col = c('skyblue','orange', 'yellow', 'black', 'green', 'purple'), title = 'HABITAT', pch=1)
```

#Step 4-D (HW Question #4 / Prompt #4): Based on your plot, describe which (if any) of the habitats differ in community composition
```{r}
print("Based on the plot above(STEP 4-D), Foothills Woodland (highlighted 'black'), Gipps.Manna (highlighted 'orange'), Box-Ironbark (highlighted 'green') indicate similarities as their elipses shape overlap. ")

cat("\n")

print("On the other hand, River Red Gum (highlighted 'purple'), Montane Forest (highlighted 'yellow') and Mixed (highlighted 'blue') do no indicate  similarities as their elipses shape do not overlap.")
```

##End of the analysis
```{r}
print("ALL ANALSYSIS DONE. CHECK HTML OUTPUT & CHECK EXPORTED FILES IF ANY. HAVE A GREAT DAY☺︎︎")
```
