---
title: "BIOL606 WED 02/13/23 HW LEC04 Cho HTML NOTEBOOK"
author: "Eddie Cho"
date: "02/13/2023"
output: html_notebook
---

#Step0: Verify current directory
```{r}
print("The analysis started...")
print(paste("YOUR current working directory is:", getwd()))
```

#Step1: Invoke necessary packages
```{r setup, message=FALSE}
library(tidyverse)
library(ggfortify)
library(multcomp)
library(ggplot2)
library(car)

print("Done. All packages loaded")
```

#Step2 (Question 1 / Prompt 1): Import lecture5_NEWcucumberdamage.csv
```{r}
new_cucumberdamage <- read.csv("lecture5_NEWcucumberdamage.csv")
print("Done. All Data loaded and saved as dataframe")

head(new_cucumberdamage) #Display first six rows data
dim(new_cucumberdamage) #check dimensions of data
```

#Step 3 (Question 2 / Prompt 2): If you're going to analyze totalfruits again, is this dataset balanced with regards to  pollination & herbivory? What about pollination & leaf damage?
```{r}
#To check the "balance-ness" of data, table() function is used.

# The following website explains more: https://www.projectpro.io/recipes/is-table-function-r-useful
##(from the website) "Table function R performs a tabulation of categorical variable and gives its frequency as output. #It is further useful to create conditional frequency table and Proportinal frequency table".

# HW Question 1 / Prompt 1 asks to do separate checks on: pollination & herbivory AND pollination & leaf damage

#table_herb checks categorical variables on pollination & herbivory
table_herb <- table(new_cucumberdamage$herbivory, new_cucumberdamage$pollination)
print(table_herb)

print("As shown above, categorical variables on pollination & herbivor (first table) looks balanced. The data contains NA, but it seems that table() function excluded. Therefore, it will be assumed that categorical variables on pollination & herbivor are based on the table. FYI, Balanced designs = equal sample size in all treatment combinations")


#table_leaf checks categorical variables on pollination & leafdamage
table_leaf <- table(new_cucumberdamage$pollination, new_cucumberdamage$leafdamage)
print(table_leaf)

print("As shown above, categorical variables on pollination & leaf damage (second table) clearly indicate it is unbalanced. FYI, Balanced designs = equal sample size in all treatment combinations ")
```


#Step 4 (Question 3 / Prompt 3) Pick one of these pairs of IVs (pollination & herbivory, or pollination & leaf  damage). Determine if these variables affect totalfruits, including an interaction, and  going through the same steps you did last time to fit the model
```{r}
#I picked pollination & leaf damage as it is unbalanced, which seems to be more prevalent in biological data
#I picked pollination & leaf damage as it is unbalanced, which seems to be more prevalent in biological data

print("Upon visually inspecting data, the dataset in unbalanced because it contains N/As")

#pollination & hervibory as model 2
model_1_a <- lm(totalfruits ~  leafdamage * herbivory, data=new_cucumberdamage)
autoplot(model_1_a)
anova(model_1_a)
summary(model_1_a)

model_1_b <- glm(totalfruits ~  leafdamage * herbivory, data=new_cucumberdamage, family=poisson)
autoplot(model_1_b)
anova(model_1_b)
summary(model_1_b)

model_1_c <- glm(totalfruits ~  leafdamage * herbivory, data=new_cucumberdamage, family=quasipoisson)
autoplot(model_1_c)
anova(model_1_c)
summary(model_1_c)

#pollination & leaf damage as model 1
model_2_a <- lm(totalfruits ~  leafdamage * pollination, data=new_cucumberdamage)
autoplot(model_2_a)

#family = poisson
model_2_b <- glm(totalfruits ~  leafdamage * pollination, data=new_cucumberdamage, family=poisson)
autoplot(model_2_b)
anova(model_2_b)
summary(model_2_b)

#family = quasi poisson
model_2_c <- glm(totalfruits ~  leafdamage * pollination, data=new_cucumberdamage, family=quasipoisson)
autoplot(model_2_c)
anova(model_2_c)
summary(model_2_c)
```
#model 3: leafdamage * herbivory
```{r}
options(contrasts = c("contr.sum","contr.poly")) #change the contrasts settings
model_3_lh <- lm(totalfruits ~ leafdamage * herbivory, data=new_cucumberdamage ) #then refit the model under the new settings
Anova(model_3_lh,type=3) #or Anova(model,type="III")
options(contrasts = c("contr.treatment","contr.poly")) #switch them back
```
#Type 1: leafdamage * herbivory
```{r}
#Type 1
options(contrasts = c("contr.sum","contr.poly")) #change the contrasts settings
model_1_lh <- lm(totalfruits ~ leafdamage * herbivory, data=new_cucumberdamage ) #then refit the model under the new settings

anova(model_1_lh) #or Anova(model,type="III")
options(contrasts = c("contr.treatment","contr.poly")) #switch them back
```

#Type 3 for leafdamage *pollination
```{r}
#Type 3
options(contrasts = c("contr.sum","contr.poly")) #change the contrasts settings
model_3_lp <- lm(totalfruits ~ leafdamage *pollination, data=new_cucumberdamage ) #then refit the model under the new settings

Anova(model_3_lp,type=3) #or Anova(model,type="III")
options(contrasts = c("contr.treatment","contr.poly")) #switch them back
```
#Type 1: for leafdamage *pollination
```{r}
#Type 1
options(contrasts = c("contr.sum","contr.poly")) #change the contrasts settings
model_1 <- lm(totalfruits ~ leafdamage *pollination, data=new_cucumberdamage ) #then refit the model under the new settings

anova(model_1) #or Anova(model,type="III")
options(contrasts = c("contr.treatment","contr.poly")) #switch them back
```
