---
title: "BIOL606 MON 02/06/23 HW LEC05 Cho HTML NOTEBOOK"
author: "Eddie Cho"
date: "02/06/2023"
output: html_notebook
---

#Step0: Verify current directory
```{r}
print("The analysis started...")
print(paste("YOUR current working directory is:", getwd() ) )
```

#Step1: Invoke necessary packages
```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
print("Done. All libraries loaded")
```

#Step2: load data and save as dataframe
```{r}
cdata <- read.csv("lecture3_cucumberdamage.csv")
flowercounts <- read.csv("lecture3_flowercounts.csv")

head(cdata)
head(flowercounts)
```

#Step 3: manipulate data
##PLEATE NOTE: codes below were provide by the professor (LEC5 slides)
##PLEATE NOTE: codes below were provide by the professor (LEC5 slides)
```{r}
#save data as.factor
cdata$plant <- as.factor(cdata$plant)

#save data as.factor
flowercounts$plant <- as.factor(flowercounts$plant)

#summarize data then save as df
flowertotals <- flowercounts %>% group_by(plant) %>%summarise(Fs = sum(F, na.rm=T), Ms = sum(M, na.rm=T))
head(flowertotals)

#left join data
cfdata <- left_join(cdata, flowertotals, by = "plant")
head(cfdata)

#create ggplot
male_plot <- ggplot(cfdata, aes(x = num_leaves, y = Ms)) + 
             geom_point(method='lm', se=TRUE)
print("male_plot")
```
#Step 4 (Question #1/Prompt #1): Form a hypothesis, being specific about it.Is total male flowers affected by number of leaves AND pollination treatment (in a single linear model)?
```{r}
print("Based on context given on lecture slide 05, the following is the hypothesis: thr higher number of leaves and higher pollination treatment by hand, the higher total male flowers will occur")
```

#Step 5 (QUESTION #2/Prompt #2): Visualize the data in a plot
```{r}
initial_plot <- ggplot(data = drop_na(cfdata), aes(x = num_leaves, y = Ms, colour=pollination )) +
                geom_point() + geom_smooth(method ='lm', se=TRUE) + 
                labs(title = 'Initial plot of number of leaves AND total Male flowers')
print(initial_plot)
```
#Step 6 (QUESTION #3/Prompt #3): Build a model using lm().
```{r}
#Do LM 
q3_model <- lm(Ms ~ num_leaves *pollination, data= cfdata)
print(q3_model)
```
#Step 7 (QUESTION #4/Prompt #4): Assess the model
```{r}
#Do ANOVA
anova(q3_model)

#Do summary
summary(q3_model)

#create four plots
plot(q3_model)
```

#Step 8 (QUESTION #5/Prompt #5): Interpret the model, particularly in light of your hypothesis.
```{r}
anova(q3_model)
print("As shown above, given p-value  of num_leaves:pollination is larger than 0.05 (The computed p-value  of num_leaves:pollination is 0.3801), it fails to reject the null hypothesis. In other words, it is not siginificant")
```


#Step 9 (QUESTION #6/Prompt #6):Plot your results the way you would 		want it to look in a journal (or on grandma's refrigerator)
```{r}
improved_plot <- ggplot(data = drop_na(cfdata), aes(x = num_leaves, y = Ms, colour=pollination )) +
                geom_point() + geom_smooth(method ='lm', se=TRUE) + 
                labs(title = 'Improved plot of number of leaves AND total Male flowers') +
                scale_fill_discrete(name = "Pollination Status ") + 
                xlab("Number of Leaves")  + 
                ylab("Total Male Flowers")
print(improved_plot)
```


##TEXTBOOK CH7
##TEXTBOOK CH7
```{r}
#read data
soay <- read.csv("soaysheepfitness.csv")
print("Done. Loaded data and saved as dataframe")

#Peek at data
glimpse(soay)

#plot data
first_plot <- ggplot(soay, aes(x = body.size, y = fitness)) +
              geom_point() +
              geom_smooth(method = "lm", se = FALSE) +
              geom_smooth(span = 1, colour = "red", se = FALSE) +
              xlab("Body mass (kg)") + ylab("Lifetime fitness")

#Display first plot
print(first_plot)

#build a GLM
soay.glm_1 <- glm(fitness ~ body.size, data = soay,family = poisson)
print(soay.glm)

#another way to do GLM
soay.glm_2 <- glm(fitness ~ body.size, data = soay, family = poisson(link = log))

#Do ANOVA
anova(soay.glm)

#Do ANOVA with chi distribution (NOTE: this is NOT chi square test)
anova(soay.glm, test = "Chisq")

#Do summary
summary(soay.glm)

#set MIN & MAX parameters
min.size <- min(soay$body.size)
max.size <- max(soay$body.size)

# make the new.x values
new.x <- expand.grid(body.size = seq(min.size, max.size, length=1000))

#generate fits and standard errors
new.y = predict(soay.glm, newdata = new.x, se.fit = TRUE)
new.y = data.frame(new.y)

#preview new.y
head(new.y)

#create DF called add.These which contains new.x and new.y
addThese <- data.frame(new.x, new.y)
addThese <- rename(addThese, fitness = fit)

#preview addThese
head(addThese)

#Use mutate to add a new column
addThese <- mutate(addThese, lwr = fitness - 1.96 * se.fit, upr = fitness + 1.96 * se.fit)

#create ggplot
second_plot <- ggplot(soay, aes(x = body.size, y = fitness)) +
               geom_point(size = 3, alpha = 0.5) +
               geom_smooth(data = addThese, aes(ymin = lwr, ymax = upr), stat = 'identity') +
               theme_bw()
print(second_plot)

print("CH7 coding all done.")
```

```{r}
print("ALL ANALSYSIS DONE. CHECK HTML OUTPUT & CHECK EXPORTED FILES IF ANY. HAVE A GREAT DAY.")
```
