---
title: "BIOL606 WED 02/08/23 HW LEC06 Cho HTML NOTEBOOK"
author: "Eddie Cho"
date: "02/08/2023"
output: html_notebook
---

#Step 0: Verify current directory
```{r}
print("The analysis started...")
print(paste("YOUR current working directory is:", getwd() ) )
```

#Step 1: Invoke necessary packages
```{r setup, include=FALSE}
library(tidyverse)
library(ggfortify)
library(ggplot2)
print("Done. All libraries loaded")
```

#Step 2: Read data & save as dataframe
```{r}
#read data
cucumberdamage <- read.csv("lecture5_cucumberdamage.csv")
print("Done. Data loaded and saved as dataframe")

#check dimension of data
dim(cucumberdamage)

#display first 6 rows of data
head(cucumberdamage)
```

#Step 3 (Question 1/Prompt 1): Is the total number of fruits produced affected by herbivore damage to leaves (i.e., "leafdamage") and hand-pollination treatments?
```{r}
print("To find out, whether the total number of fruits produced is affected by herbivore damage to leaves, linear model is plotted")

#Do basic lm model 
model_1 <- lm(totalfruits ~  leafdamage * pollination, data=cucumberdamage)
autoplot(model_1)

anova(model_1)
plot(model_1)

#Do glm model
model_2 <- glm(totalfruits ~ leafdamage * pollination, data=cucumberdamage, family=poisson())
autoplot(model_2)

anova(model_2)
plot(model_2)

print("Based on the computed lm table, leafdamage:pollination is larger than 0.05, (exact value is 0.5909), it seems that it is not siginificant enough to rejct the null hypothesis.")
```

#Step 4 (Question 2/Prompt 2): Make a figure of the results, with separate lines for the two pollination treatments
```{r}
first_plot <- ggplot(cucumberdamage, aes(x=totalfruits, y=leafdamage, color=pollination)) +  
              geom_point() + 
              theme_bw() + 
              geom_smooth(method = 'glm', se=FALSE) +
              xlab("Total Fruits")  +  #add x axis label
              ylab("Leaf Damage") + #add y axis label
              labs(title = 'First plot: Total Fruits vs. Leaf Damage colored by Pollination')+ #add title
              coord_cartesian(ylim = c(0,40)) #adjust y axis ticks marks (0 to 40)
#Display plot
print(first_plot)

#Save plot
ggsave("First_Plot.png", plot = first_plot)
```

#Step 5-A (Question 3/Prompt 3): Repeat steps 1 and 2, but instead of "leafdamage," use the categorical treatment variable "herbivory". Think about what the figure will look like for this scenario
```{r}
#do lm and glm using another categorical treatment variable, herbivory

#LM
model_3_a <- lm(totalfruits ~ herbivory * pollination, data=cucumberdamage)
autoplot(model_3_a)
anova(model_3_a)
plot(model_3_a)

#GLM
model_3_b <- glm(totalfruits ~ herbivory * pollination, data=cucumberdamage, family=poisson())
autoplot(model_3_b)
anova(model_3_b)
plot(model_3_b)

print("Based on the computed lm table, herbivory:pollination is larger than 0.05, (exact value is 0.4763), it seems that it is not siginificant enough to rejct the null hypothesis.")
```

#Step 5-B (Question 3/Prompt 3): Repeat steps 1 and 2, but instead of "leafdamage," use the categorical treatment variable "herbivory". Think about what the figure will look like for this scenario
```{r}
print("For this scenario, the figure will be box plot")

model_3_b <- ggplot(cucumberdamage, aes(x=totalfruits, y=herbivory, color=pollination)) +  
             geom_boxplot() + 
             theme_bw() +
             coord_flip() + #Flip box plot
             xlab("Total Fruits")  +  #add x axis label
             ylab("Herbivory") + #add y axis label
             labs(title = 'Box Plot of Total Fruits vs. Herbivory colored by Pollination') #add title

#Display plot
print(model_3_b)

#Save plot
ggsave("Box_Plot.png", plot = model_3_b)
```
##INCLASS
```{r}
cucumberdamage <- read.csv("lecture5_cucumberdamage.csv")
icuke1 <- glm(totalfruits ~ leafdamage * pollination, data = cucumberdamage, family=quasipoisson)
anova(icuke1, test="F")

icuke2 <- glm(totalfruits ~ pollination * leafdamage , data = cucumberdamage, family=quasipoisson)
anova(icuke2, test="F")
```

```{r}
print("ALL ANALSYSIS DONE. CHECK HTML OUTPUT & CHECK EXPORTED FILES IF ANY. HAVE A GREAT DAY.")
```