---
title: "BIOL606 WED 4/12/23 HW LEC14 Cho HTML NOTEBOOK"
author: "Eddie Cho"
date: "04/12/2023"
output:
  html_document:
    df_print: paged
---

#Step 0: Verify current directory
```{r}
print("The analysis started...")
print(paste("YOUR current working directory is:", getwd()))
```

#Step 1: Invoke necessary libraries/packages
```{r setup, message=FALSE}

library(dplyr)
library(ggplot2)
library(tidyverse)
library(ggfortify)

print("Done. All packages loaded")
```

#Step 2: Import data
```{r}
bump <-read.table("bumpus.dat.txt", header=TRUE)
print("Data loaded and saved as dataframe called 'bump' ")

head(bump)
dim(bump)
```

#Step 3 (HW Question #1 / Prompt #1): Produce a pairs plot of the various bird traits
```{r}
pairs(bump[,c(-1)], lower.panel=NULL)
```

#Step 4 (HW Question #2 / Prompt #2): Fit a PCA, using centered and scaled trait data. Show the summary output for the PCA
```{r}
pca_plot <- prcomp(bump[,c(-1)], center=TRUE, scale.=TRUE)

summary(pca_plot)

screeplot(pca_plot)
```

#Step 5 (HW Question #3 / Prompt #3): Produce a biplot using the ggfortify package autoplot() function, showing the birds colored by survival, and arrows showing the variables in the PCA
```{r}
#Create basic bi plot
bi_plot_basic <- biplot(pca_plot, cex= c(1,1.3), xlab="PC_1", ylab="PC_2")
```

#Step 6 (HW Question #4 / Prompt #4):  Produce a biplot using the ggfortify package autoplot() function, showing the birds colored by survival, and arrows showing the variables in the PCA
```{r}

#Create advanced bi plot
bi_plot_advanced <- autoplot(pca_plot, data = bump, colour = 'survive', loadings=TRUE, size = 3, loadings.label = TRUE, loadings.label.size=5,label=TRUE,label.label = 'survive', label.vjust = -0.4)

print(bi_plot_advanced)
```


##End of the analysis
```{r}
print("ALL ANALSYSIS DONE. CHECK HTML OUTPUT & CHECK EXPORTED FILES IF ANY. HAVE A GREAT DAY☺︎︎")
```
